---
title: "MSDS 6306 Case Study 2"
author: "by Inderbir Dhillon"
date: "4/14/2020"
output: html_document
---


## Analysis of Employee Retention by DDSAnalytics
### A video presentation of this data can be found at: https://www.youtube.com/watch?v=s4M4MJ1SODw&feature=youtu.be

### In this report, we are trying to predict employee retention based on a number of features.


```{r}
# load in the libraries
library(class)
library(caret)
library(e1071)
library(GGally)
library(corrplot)
library(corrgram)
library(naniar)
library(gridExtra)
library(tidyverse)
```

```{r}
attr = read_csv(
  "./data/CaseStudy2-data.csv",
  col_types = cols(
    Attrition = col_character(),
    BusinessTravel = col_factor(),
    Department = col_factor(),
    EducationField = col_factor(),
    Gender = col_factor(),
    JobRole = col_factor(),
    MaritalStatus = col_factor(),
    OverTime = col_factor(),
    Age = col_integer(),
    DailyRate = col_integer(),
    DistanceFromHome = col_integer(),
    EmployeeNumber = col_integer(),
    HourlyRate = col_integer(),
    MonthlyIncome = col_integer(),
    MonthlyRate = col_integer(),
    NumCompaniesWorked = col_integer(),
    PercentSalaryHike = col_integer(),
    TotalWorkingYears = col_integer(),
    TrainingTimesLastYear = col_integer(),
    YearsAtCompany = col_integer(),
    YearsInCurrentRole = col_integer(),
    YearsSinceLastPromotion = col_integer(),
    YearsWithCurrManager = col_integer(),
    Education = col_integer(),
    EnvironmentSatisfaction = col_integer(),
    JobInvolvement = col_integer(),
    JobLevel = col_integer(),
    JobSatisfaction = col_integer(),
    PerformanceRating = col_integer(),
    RelationshipSatisfaction = col_integer(),
    StockOptionLevel = col_integer(),
    WorkLifeBalance = col_integer()
  )
)
attr = attr %>% select(-one_of( c("ID", "EmployeeCount", "Over18", "StandardHours")))
```

### We see that there are no missing values.

```{r}
gg_miss_var(attr)
```


```{r}
continuous_features = attr %>% select(
  Age,
  DailyRate,
  DistanceFromHome,
  EmployeeNumber,
  HourlyRate,
  MonthlyIncome,
  MonthlyRate,
  NumCompaniesWorked,
  PercentSalaryHike,
  TotalWorkingYears,
  TrainingTimesLastYear,
  YearsAtCompany,
  YearsInCurrentRole,
  YearsSinceLastPromotion,
  YearsWithCurrManager,
  Education,
  EnvironmentSatisfaction,
  JobInvolvement,
  JobLevel,
  JobSatisfaction,
  PerformanceRating,
  RelationshipSatisfaction,
  StockOptionLevel,
  WorkLifeBalance
)

factors = attr %>% select(
  BusinessTravel,
  Department,
  EducationField,
  Gender,
  JobRole,
  MaritalStatus,
  OverTime
)
```

### First we look at the collinearity of the continuous variables and notice that manu of the year realted vairables are correlated


```{r}
corrgram(continuous_features)
```


```{r}
uncorrelated_features = c(
  'DailyRate', 
  'DistanceFromHome', 
  'EmployeeNumber', 
  'HourlyRate', 
  'MonthlyRate', 
  'TrainingTimesLastYear', 
  'EnvironmentSatisfaction', 
  'JobInvolvement', 
  'JobSatisfaction', 
  'RelationshipSatisfaction', 
  'StockOptionLevel', 
  'WorkLifeBalance'
)
feat1  = attr %>% select(Attrition, YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager)
ggpairs(feat1, mapping=ggplot2::aes(color = Attrition), progress = FALSE)
```

### Breaking down the years features, we do see there is some realation between these features and attrition. This relationship is easier to see on a log scale.

```{r}
attr$LogYearsAtCompany = log(attr$YearsAtCompany + 1)
attr$LogYearsInCurrentRole = log(attr$YearsInCurrentRole + 1)
attr$LogYearsSinceLastPromotion = log(attr$YearsSinceLastPromotion + 1)
attr$LogYearsWithCurrManager = log(attr$YearsWithCurrManager + 1)
feat2 = attr %>% select(Attrition, LogYearsAtCompany, LogYearsInCurrentRole, LogYearsSinceLastPromotion, LogYearsWithCurrManager)
ggpairs(feat2, mapping=ggplot2::aes(color = Attrition), progress = FALSE)
```
```{r}
attr$YearsBeforeCompany = attr$TotalWorkingYears - attr$YearsAtCompany
attr$LogTotalWorkingYears = log(attr$TotalWorkingYears + 1)
feat3  = attr %>% select(Attrition, Age, MonthlyIncome, YearsBeforeCompany, JobLevel)
ggpairs(feat3, mapping=ggplot2::aes(color = Attrition), progress = FALSE)
```

### Age and income also look quite important when looking at attrition

```{r}
attr$LogMonthlyIncome = log(attr$MonthlyIncome + 1)
attr$LogYearsBeforeCompany= log(attr$YearsBeforeCompany + 1)
feat3  = attr %>% select(Attrition, Age, LogMonthlyIncome, LogYearsBeforeCompany, JobLevel)
ggpairs(feat3, mapping=ggplot2::aes(color = Attrition), progress = FALSE)
```

```{r}
feat4  = attr %>% select(Attrition, PercentSalaryHike, PerformanceRating, NumCompaniesWorked, Education, YearsBeforeCompany)
ggpairs(feat4, mapping=ggplot2::aes(color = Attrition), progress = FALSE)
```
```{r}
attr$LogNumCompaniesWorked = log(attr$NumCompaniesWorked + 1)
attr$LogYearsBeforeCompany = log(attr$YearsBeforeCompany + 1)
feat5  = attr %>% select(Attrition, PercentSalaryHike, PerformanceRating, LogNumCompaniesWorked, Education, LogYearsBeforeCompany)
ggpairs(feat5, mapping=ggplot2::aes(color = Attrition), progress = FALSE)
```

### Education, Number of Companies, and Years before company also look like they could add some information.

```{r}
attr$LogDistanceFromHome = log(attr$DistanceFromHome + 1)
feat6  = attr %>% select(Attrition, DailyRate, HourlyRate, MonthlyRate, LogDistanceFromHome)
ggpairs(feat6, mapping=ggplot2::aes(color = Attrition), progress = FALSE)
```

```{r}
feat6  = attr %>% select(Attrition, EmployeeNumber, TrainingTimesLastYear, StockOptionLevel, WorkLifeBalance)
ggpairs(feat6, mapping=ggplot2::aes(color = Attrition), progress = FALSE)
```

### Stock level and work/life balalance also look like they could add some influnece.

```{r}
feat7  = attr %>% select(Attrition, EnvironmentSatisfaction, JobInvolvement, RelationshipSatisfaction, JobSatisfaction)
ggpairs(feat7, mapping=ggplot2::aes(color = Attrition), progress = FALSE)
```

### We can see if taking a single measure accross satiscaction field ties the inforation together 

```{r}
attr$pressurePoint = pmin(
  attr$JobSatisfaction, 
  attr$EnvironmentSatisfaction, 
  attr$RelationshipSatisfaction
)
attr$TotalSatisfaction = pmax(
  attr$JobSatisfaction, 
  attr$EnvironmentSatisfaction, 
  attr$RelationshipSatisfaction
)
attr$pressurePointFull = pmin(
  attr$JobSatisfaction, 
  attr$EnvironmentSatisfaction, 
  attr$RelationshipSatisfaction, 
  attr$JobInvolvement, 
  attr$WorkLifeBalance
)
attr$TotalSatisfactionFull = pmax(
  attr$JobSatisfaction, 
  attr$EnvironmentSatisfaction, 
  attr$RelationshipSatisfaction, 
  attr$JobInvolvement, 
  attr$WorkLifeBalance
)
feat7  = attr %>% select(Attrition, pressurePoint, TotalSatisfaction, pressurePointFull, TotalSatisfactionFull)
ggpairs(feat7, mapping=ggplot2::aes(color = Attrition), progress = FALSE)
```




```{r}
cat_feats = attr %>% select(    
    Attrition,
    BusinessTravel,
    Department,
    EducationField,
    Gender,
    JobRole,
    MaritalStatus,
    OverTime
)
ggplot(data = cat_feats, mapping = aes(x=JobRole, fill = Attrition)) + geom_bar(position="fill") + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
tr = ggplot(data = cat_feats, mapping = aes(x=BusinessTravel, fill = Attrition)) + geom_bar(position="fill")
dep = ggplot(data = cat_feats, mapping = aes(x=Department, fill = Attrition)) + geom_bar(position="fill")
ge = ggplot(data = cat_feats, mapping = aes(x=Gender, fill = Attrition)) + geom_bar(position="fill")
ma = ggplot(data = cat_feats, mapping = aes(x=MaritalStatus, fill = Attrition)) + geom_bar(position="fill")
ov = ggplot(data = cat_feats, mapping = aes(x=OverTime, fill = Attrition)) + geom_bar(position="fill")
grid.arrange(tr, dep, ge, ma, ov)

```

```{r}
ggplot(data = cat_feats, mapping = aes(x=EducationField, fill = Attrition)) + geom_bar(position="fill") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



```{r}
ggplot(data = attr, mapping = aes(x=JobRole, y=LogMonthlyIncome)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### It is interesting to note that the job levels with the highest salary also have the lowest job satisfcation.

```{r}
ggplot(data = attr, mapping = aes(x=JobRole, y= JobSatisfaction )) + geom_boxplot() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


### Stock option has more of a bathtub shaped attrition curve.

```{r}
ggplot(data = attr, mapping = aes(x=StockOptionLevel, fill = Attrition)) + geom_bar(position="fill") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



### Drop the columns that are are not useful or those that have been replaced by their log version

```{r}
drop_cols = c(
  'YearsAtCompany',
  'YearsInCurrentRole',
  'LogYearsSinceLastPromotion',
  'YearsWithCurrManager',
  'YearsBeforeCompany',
  'TotalWorkingYears',
  'NumCompaniesWorked',
  'MonthlyIncome',
  'DistanceFromHome',
  'Gender',
  'EmployeeNumber'
)
attr = attr %>% select(-one_of(drop_cols))
```



```{r}
attr$Attrition <- attr$Attrition %>% recode('Yes'=1,'No'= 0)
attr$BusinessTravel <- attr$BusinessTravel %>% recode('Travel_Frequently'=2,'Travel_Rarely'=1,'Non-Travel'= 0)
attr$Department <- attr$Department %>% recode('Sales'=2,'Human Resources'=1,'Research & Development'= 0)
attr$EducationField <- attr$EducationField %>% recode('Medical'=0,'Life Sciences'=1,'Other'=2, 'Marketing'=3, 'Technical Degree'=4, 'Human Resources'=5)
attr$OverTime <- attr$OverTime %>% recode('Yes'=1,'No'= 0)
attr$MaritalStatus <- attr$MaritalStatus %>% recode('Divorced'=2,'Married'=1,'Single'= 0)
attr$JobRole <- attr$JobRole %>% recode(
  'Sales Representative'=0, 
  'Human Resources'=1, 
  'Laboratory Technician'=2, 
  'Sales Executive'=3, 
  'Research Scientist'=4, 
  'Healthcare Representative'=5, 
  'Manager'=6,
  'Research Director'=7,
  'Manufacturing Director'=7
)
```


### Split into train and test sets to evalute the performance of the KNN model

```{r}
splitPerc = .70
trainIndices = sample(1:dim(attr)[1],round(splitPerc * dim(attr)[1]))
train = attr[trainIndices,]
test = attr[-trainIndices,]
train_labels = train$Attrition
test_labels = test$Attrition

test = test %>% select(-Attrition)
train = train %>% select(-Attrition)

accs = data.frame(sens = numeric(30), spec = numeric(30), k = numeric(30))

for(i in 1:30)
{
  classifications = knn(train, test, train_labels, prob = TRUE, k = i)
  table(test_labels, classifications)
  CM = confusionMatrix(table(test_labels, classifications))
  accs$sens[i] = CM$byClass['Sensitivity']
  accs$spec[i] = CM$byClass['Specificity']
  accs$k[i] = i
}

accs[is.na(accs)] = 0
print(max(accs$spec))
```

### Taking a subset of the features produces a mch better model.

```{r}
set.seed(742)

exp_features = attr %>% select(
  Attrition,
  JobSatisfaction,
  MaritalStatus,
  JobRole,
  StockOptionLevel,
  LogMonthlyIncome,
  LogYearsInCurrentRole,
)

splitPerc = .70
trainIndices = sample(1:dim(exp_features)[1],round(splitPerc * dim(exp_features)[1]))
train = exp_features[trainIndices,]
test = exp_features[-trainIndices,]
train_labels = train$Attrition
test_labels = test$Attrition

test = test %>% select(-Attrition)
train = train %>% select(-Attrition)

accs = data.frame(sens = numeric(40), spec = numeric(40), k = numeric(40))

for(i in 1:40)
{
  classifications = knn(train, test, train_labels, prob = TRUE, k = i)
  table(test_labels, classifications)
  CM = confusionMatrix(table(test_labels, classifications))
  accs$sens[i] = CM$byClass['Sensitivity']
  accs$spec[i] = CM$byClass['Specificity']
  accs$k[i] = i
}

print(accs)
```

```{r}
comp_noattr = read_csv("./data/CaseStudy2CompSet No Attrition.csv",
  col_types = cols(
    JobRole = col_factor(),
    MaritalStatus = col_factor(),
    MonthlyIncome = col_integer(),
    YearsInCurrentRole = col_integer(),
    JobSatisfaction = col_integer(),
    StockOptionLevel = col_integer()
  )
)
comp_noattr$LogYearsInCurrentRole = log(comp_noattr$YearsInCurrentRole + 1)
comp_noattr$LogMonthlyIncome = log(comp_noattr$MonthlyIncome + 1)

comp_noattr$BusinessTravel <- comp_noattr$BusinessTravel %>% recode('Travel_Frequently'=2,'Travel_Rarely'=1,'Non-Travel'= 0)
comp_noattr$Department <- comp_noattr$Department %>% recode('Sales'=2,'Human Resources'=1,'Research & Development'= 0)
comp_noattr$EducationField <- comp_noattr$EducationField %>% recode('Medical'=0,'Life Sciences'=1,'Other'=2, 'Marketing'=3, 'Technical Degree'=4, 'Human Resources'=5)
comp_noattr$OverTime <- comp_noattr$OverTime %>% recode('Yes'=1,'No'= 0)
comp_noattr$MaritalStatus <- comp_noattr$MaritalStatus %>% recode('Divorced'=2,'Married'=1,'Single'= 0)
comp_noattr$JobRole <- comp_noattr$JobRole %>% recode(
  'Sales Representative'=0, 
  'Human Resources'=1, 
  'Laboratory Technician'=2, 
  'Sales Executive'=3, 
  'Research Scientist'=4, 
  'Healthcare Representative'=5, 
  'Manager'=6,
  'Research Director'=7,
  'Manufacturing Director'=7
)

comp_noattr_model = comp_noattr %>% select(
  JobSatisfaction,
  MaritalStatus,
  JobRole,
  StockOptionLevel,
  LogMonthlyIncome,
  LogYearsInCurrentRole,
)

comp_classifications = knn(train, comp_noattr_model, train_labels, k = 10)
```

```{r}
comp_noattr$Attribution =  comp_classifications
comp_noattr = comp_noattr %>% select(ID, Attribution)
write_csv(comp_noattr, 'Case2PredictionsDhillon Attrition.csv')

```



```{r}
income_feats = attr %>% mutate(EducationField = factor(EducationField), JobRole = factor(JobRole))
income_feats = income_feats %>% select(Age, JobRole, EducationField, LogTotalWorkingYears, LogMonthlyIncome)

splitPerc = .70
trainIndices = sample(1:dim(exp_features)[1],round(splitPerc * dim(exp_features)[1]))
train = income_feats[trainIndices,]
test = income_feats[-trainIndices,]

income_model <- lm(LogMonthlyIncome ~ Age + JobRole + LogTotalWorkingYears, data=train)
summary(income_model)
```


```{r}
comp_noIncome = read_csv("./data/CaseStudy2CompSet No Salary.csv",
  col_types = cols(
    JobRole = col_factor(),
    TotalWorkingYears = col_integer(),
    Age = col_integer()
  )
)
comp_noIncome$JobRole <- comp_noIncome$JobRole %>% recode(
  'Sales Representative'=0, 
  'Human Resources'=1, 
  'Laboratory Technician'=2, 
  'Sales Executive'=3, 
  'Research Scientist'=4, 
  'Healthcare Representative'=5, 
  'Manager'=6,
  'Research Director'=7,
  'Manufacturing Director'=7
)
comp_noIncome = comp_noIncome %>% mutate(JobRole = factor(JobRole))

comp_noIncome$LogTotalWorkingYears = log(comp_noIncome$TotalWorkingYears + 1)
comp_noIncome_model = comp_noIncome %>% select(Age, JobRole, LogTotalWorkingYears)
pred_salary = predict(income_model, comp_noIncome_model)
comp_noIncome$LogMonthlyIncome = pred_salary
comp_noIncome$MonthlyIncome = exp(comp_noIncome$LogMonthlyIncome) - 1
comp_noIncome = comp_noIncome %>% select(ID, MonthlyIncome)
write_csv(comp_noIncome, 'Case2PredictionsDhillon Salary.csv')
```


